Type Error Telemetry at Scale

### 2023-09-30 programming23/paper.tex
type error in range + popularity

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1\vphantom{Ap}}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1\vphantom{Ap}}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1\vphantom{Ap}}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}

@@

\paragraph{Number of Type Analysis Errors}
%\label{s:count-analysis-errors}

@@

  \newcommand{\labelbars}[1]{\begin{tabular}[t]{l@{}l} \begin{tabular}[t]{r} \mnocheck{} \\ \mnonstrict{} \\ \mstrict{} \\ ~ \end{tabular} & #1 \end{tabular}}

    \begin{tabular}[t]{cc}
    %% TODO why text blurry??!
    Type errors & Background errors \\
    \labelbars{\includegraphics[valign=M]{img/error-by-mode-te.pdf}}
    & \labelbars{\includegraphics[valign=M]{img/error-by-mode-fs.pdf}}
  \end{tabular}
  \caption{FILL}
  \label{f:error-by-mode}
\end{figure}

\Cref{f:error-by-mode} shows how the type errors and \FS{} errors
in the data are split across type analysis modes.

Most type errors come from \mnonstrict{} mode, despite it having an order
of magnitude fewer users that \mnocheck{}.
This is unsurprising because \mnocheck{} reports only syntax errors,
whereas \mnonstrict{} reports several others kinds~(detailed in the next section).
Finally, \mstrict{} mode accounts for a small fraction of type errors, despite
being more picky, because so few creators use it.

Background analysis applies the same typechecker to every module, regardless
of mode, in the hope of enabling IDE tools.
Consequently, the breakdown of \FS{} errors matches the overall usage
data~(\cref{f:dataset-overview}): almost \pct{90} for \mnocheck{},
\pct{10} for \mnonstrict{}, and \pct{1} for \mstrict{}.
Type analysis finds a large number of errors in \mnocheck{} code;
unfortunately, we have no insight into whether those errors are true
positives or whether the analysis led to useful IDE suggestions.

\FILL{} average number of FS errors is
worst in \mnonstrict{}, 49, vs 48 nocheck and 7 strict.
median is 3 NS, 1 nocheck, 1 strict.
We might optimistically think that nocheck accumulates
lots of errors, but no that does not seem to be the case,
NS and nocheck both apparently fine with latent FS errors.

@@

\begin{table}[t]
  \caption{Number of telemetry records that add, keep, or drop (remove) specific type errors from their edit range.}
  \label{t:type-error-survival}
  \begin{tabular}{l@{}r@{}r@{}rr@{}r@{}rr@{}r@{}r}
    & \zerowidth{\mnocheck{}} & & & \zerowidth{\mnonstrict{}} & & & \zerowidth{\mstrict{}} & & \\
    & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} & \rbox{Add} & \ybox{Keep} & \gbox{Drop} \\\midrule
    \code{CannotCallNonFunction} & {0} & {0} & {0} & {17} & {4} & {20} & {1} & {0} & {1} \\
    \code{CannotExtendTable} & {0} & {0} & {0} & {7} & {9} & {10} & {0} & {0} & {1} \\
    \code{CannotInferBinaryOperation} & {0} & {0} & {0} & {1} & {2} & {1} & {7} & {6} & {9} \\
    \code{CountMismatch} & {0} & {0} & {0} & {223} & {48} & {269} & {12} & {2} & {17} \\
    \code{DuplicateTypeDefinition} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {0} \\
    \code{FunctionDoesNotTakeSelf} & {0} & {0} & {0} & {4} & {6} & {5} & {0} & {0} & {0} \\
    \code{FunctionExitsWithoutReturn} & {0} & {0} & {0} & {9} & {3} & {6} & {11} & {5} & {8} \\
    \code{GenericError} & {0} & {0} & {0} & {206} & {52} & {203} & {8} & {1} & {9} \\
    \code{GenericExtraInformation} & {0} & {0} & {0} & {48} & {7} & {39} & {2} & {0} & {1} \\
    \code{IllegalRequire} & {0} & {0} & {0} & {16} & {3} & {33} & {0} & {0} & {2} \\
    \code{IncorrectGenericParamCount} & {0} & {0} & {0} & {0} & {0} & {0} & {1} & {0} & {1} \\
    \code{MissingProperties} & {0} & {0} & {0} & {11} & {3} & {11} & {6} & {4} & {6} \\
    \code{MissingUnionProperty} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
    \code{ModuleHasCyclicDependency} & {0} & {0} & {0} & {20} & {9} & {15} & {0} & {0} & {1} \\
    \code{NotATable} & {0} & {0} & {0} & {7} & {2} & {8} & {2} & {0} & {1} \\
    \code{OccursCheckFailed} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {0} & {1} \\
    \code{OnlyTablesCanHaveMethods} & {0} & {0} & {0} & {1} & {0} & {2} & {0} & {0} & {0} \\
    \code{OptionalValueAccess} & {0} & {0} & {0} & {40} & {52} & {27} & {6} & {2} & {7} \\
    \code{SyntaxError} & {8349} & {314} & {19220} & {983} & {46} & {2658} & {16} & {0} & {71} \\
    \code{TypeMismatch} & {0} & {0} & {0} & {159} & {62} & {129} & {21} & {10} & {28} \\
    \code{TypesAreUnrelated} & {0} & {0} & {0} & {1} & {0} & {0} & {0} & {0} & {1} \\
    \code{UnknownPropButGotLikeProp} & {0} & {0} & {0} & {26} & {18} & {21} & {1} & {0} & {1} \\
    \code{UnknownProperty} & {0} & {0} & {0} & {347} & {180} & {385} & {19} & {19} & {35} \\
    \code{UnknownRequire} & {0} & {0} & {0} & {91} & {39} & {85} & {12} & {3} & {4} \\
    \code{UnknownSymbol} & {0} & {0} & {0} & {2348} & {496} & {2973} & {49} & {19} & {55}
  \end{tabular}

@@

Equipped with an overall picture of the dataset and the skewed usage
of \mnocheck{}, \mnonstrict{}, and \mstrict{} modes, we are ready to explore
the type error data.
Our main focus is on the intersection between type errors
and edits to the programs.\todo{The next two sentences are in some
  sense our main RQs, right? I would call them out more explicitly and
  w/ more justification.}
When a type error highlights part of the code and the next edit range includes
that code, we want to know whether the error survives the edit.
Similarly, we want to know when an edit introduces a type error.

\Cref{t:type-error-survival} counts the number of times that types
errors overlap with the current edit range.
There are three sets of columns, corresponding to the \mnocheck{},
\mnonstrict{}, and \mstrict{} type analysis modes.
Within each set, the Add columns counts the number of telemetry records
that introduce a certain type error.
For example, seventeen of the \mnonstrict{} records add
at least one \code{CannotCallNonFunction} error to the current edit range.
The Keep columns counts records that have the same, non-zero number of type errors
before and after.
Lastly, the Remove columns count records that decrease the number of type errors.

All three counts are imprecise because our telemetry does not track the exact
identity of a type error.
Suppose a creator removes one \code{CannotCallNonFunction} error but then
immediately introduces another on a different line.
\FILL{} beware, don't say this same thing Nx throughout the paper.
\Cref{t:type-error-survival} reports this pair of edits as a Keep rather than
an Add and a Remove because we cannot distinguish two errors of the same kind.

@@

%% \paragraph{Most Type Errors are Typos}
\begin{itemize}
  \item
    \FILL{} which survive the longest, most keeps???

  \item
    \FILL{} beware counts include module switches. Why???
    Is that defensible at all?

  \item
    \FILL{} \mnonstrict{} should have no false positives,
    \mstrict{} doesn't care.
    Why so many keeps in NS??!

  \item
    One striking aspect of~\cref{t:type-error-survival} is the large number
    of \code{SyntaxError}s and the related errors \code{UnknownSymbol} and
    \code{UnknownProperty}.
    Most type errors in the data are likely due to typos.
    \Cref{s:type-error-count} discusses this point in detail.
    For now, we remark that \code{SyntaxError} is the only possible error
    in \mnocheck{} mode (by design) and observe that these errors are
    usually in the Drop column and seldom in the Keep column.\todo{So?}

  \item
    The numbers are relatively low in general, which means that
    few errors in the dataset overlap with the edit range
    The highest value in the \mstrict{} columns is 71, meaning 71 records
    out of 6,505 total~(\cref{f:dataset-overview} drop a syntax error.
    The \mnonstrict{} and \mnocheck{} columns are similar.
    %% FILL show percentages for all three modes?
    \FILL{} what does this mean? Not sure if bad, or inconclusive.

  \item
    %% Drop >= Add >= Rem ~= 27 triples
    %% other nonzero      ~= 13 triples
    When type errors overlap with the edit range,
    we would hope the number of Adds and Drops is roughly
    equal and the number of Keeps is low.
    This is indeed the case.
    In most Add/Keep/Drop triples, the Drop column has the highest
    value and the Keep column has the lowest value.
    The optimistic conclusion is that creators are fixing the code
    highlighted by errors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%  STOP!  ROUGH TEXT BELOW     %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  \item
    After the syntax errors, \code{CountMismatch},
    \code{TypeMismatch}, and \code{GenericError} are the most common
    \mnonstrict{} and \mstrict{} errors.
    \FILL{} so what.

  \item
    \FILL{} which errors break the Drop > Add > Keep ``preferred order''?

  \item
    (NOTE may want to save relative popularity for next section.)
    Some errors occur more often in \mnonstrict{} than \mstrict{}:
    \code{CannotCallNonFunction},
    \code{CannotExtendTable},
    \code{GenericExtraInformation},
    \code{FunctionDoesNotTakeSelf},
    \code{IllegalRequire},
    \code{ModuleHasCyclicDependency},
    \code{OptionalValueAccess},
    \code{UnknownPropButGotLikeProp}.
    \QALAN{} are any of these surprising, or expected from the bigger number of
    \mnonstrict{} records?

  \item
    A few errors are more common in \mstrict{} than \mnonstrict{}:
    \code{CannotInferBinaryOperation},
    \code{FunctionExitsWithoutReturn},
    \code{IncorrectGenericParamCount}.
    \FILL{} so what.

@@

\paragraph{Do Error Highlights Help?}

\FILL{} data seems to say yes.

@@

\begin{table}[t]\centering
  %% NOTE: popularity looks similar when [over all records] vs. [restricted to module switch records]
  \caption{Type error popularity for \mnonstrict{} and \mstrict{} modes (all of the type errors in \mnocheck{} mode are \code{SyntaxError}s).}
  \label{t:type-error-count}
  \begin{tabular}{ll}
    \mnonstrict{} & \mstrict{} \\
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{62.13} \\
      \code{SyntaxError} & \pct{15.42} \\
      \code{UnknownProperty} & \pct{8.28} \\
      \code{UnknownRequire} & \pct{3.13} \\
      \code{TypeMismatch} & \pct{2.44} \\
      \code{CountMismatch} & \pct{2.26} \\
      \code{OptionalValueAccess} & \pct{2.07} \\
      \code{GenericError} & \pct{2.06} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.44} \\
      \code{CannotExtendTable} & \pct{0.42} \\
      \code{GenericExtraInformation} & \pct{0.32} \\
      \code{ModuleHasCyclicDependency} & \pct{0.23} \\
      \code{IllegalRequire} & \pct{0.20} \\
      \code{NotATable} & \pct{0.15} \\
      \code{CannotCallNonFunction} & \pct{0.15} \\
      \code{MissingProperties} & \pct{0.09} \\
      \code{FunctionExitsWithoutReturn} & \pct{0.07} \\
      \code{FunctionDoesNotTakeSelf} & \pct{0.07} \\
      \code{MissingUnionProperty} & \pct{0.02} \\
      \code{CannotInferBinaryOperation} & \pct{0.02} \\
      \code{OnlyTablesCanHaveMethods} & \pct{0.01} \\
      \code{DuplicateTypeDefinition} & {<\pct{0.01}} \\
      \code{TypesAreUnrelated} & {<\pct{0.01}} \\
    \end{tabular}
        &
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{23.97} \\
      \code{TypeMismatch} & \pct{20.46} \\
      \code{UnknownProperty} & \pct{18.88} \\
      \code{SyntaxError} & \pct{9.31} \\
      \code{CannotInferBinaryOperation} & \pct{6.94} \\
      \code{MissingProperties} & \pct{4.04} \\
      \code{CountMismatch} & \pct{2.99} \\
      \code{OptionalValueAccess} & \pct{2.99} \\
      \code{FunctionExitsWithoutReturn} & \pct{2.99} \\
      \code{UnknownRequire} & \pct{2.37} \\
      \code{GenericError} & \pct{2.28} \\
      \code{NotATable} & \pct{1.32} \\
      \code{GenericExtraInformation} & \pct{0.35} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.26} \\
      \code{IncorrectGenericParamCount} & \pct{0.18} \\
      \code{CannotCallNonFunction} & \pct{0.18} \\
      \code{IllegalRequire} & \pct{0.18} \\
      \code{ModuleHasCyclicDependency} & \pct{0.09} \\
      \code{CannotExtendTable} & \pct{0.09} \\
      \code{OccursCheckFailed} & \pct{0.09} \\
      \code{TypesAreUnrelated} & \pct{0.09} \\
    \end{tabular}
  \end{tabular}
\end{table}

@@

\Cref{t:type-error-count} reports the percent of
type errors that overlap with the edit range.
This table is based on the same data as the previous table~(\cref{t:type-error-survival}),
but focuses on the overall popularity of type errors
rather than their relation to edits.
The left column in each row names one kind of type error
that overlaps with an edit range.
The right column shows what percentage of all overlapping
errors the current error took up.

There are two subtables, for \mnonstrict{} and \mstrict{} mode.
The table does not report \mnocheck{} because all its type errors are
by definition \code{SyntaxError}s.

@@

\begin{itemize}
  \item
    As we noted above, the most popular errors are syntax errors
    (\FILL{} \pct{0} \mnonstrict{}, \pct{0} \mstrict{})..
    This is undoubtedly due to the stochiastic nature
    of our telemetry, which has no way to avoid
    reporting edits that happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will
    get type analysis to find errors like \code{UnknownSymbol(buck)}
    and \code{UnknownProperty(bucket.c)}.
    These are transient issues, but telemetry may catch them.

  \item
    \FILL{} huge syntax error gap between S nonstrict.

  \item
    \code{TypeMismatch} is far more common in \mstrict{} (\pct{20.46})
    than \mnonstrict{} (\pct{2.44}) because \mstrict{} mode
    has additional type constraints.
    Several other errors have higher \mstrict{} percentages for the
    same reason, for example, \code{CannotInferBinaryOperation}.

>>>>

\newcommand{\code}[1]{\texttt{#1}}
\newcommand{\FILL}{\textbf{FILL}}
\newcommand{\dotscale}[1]{\scalebox{0.72}{#1}}
\newcommand{\wideas}[2]{\makebox[\widthof{#2}][l]{#1}}
\newcommand{\twoline}[2]{\parbox[s]{1.4cm}{\flushleft#1\newline#2}}
\newcommand{\chkYes}{\dotscale{\CIRCLE}}
\newcommand{\chkMaybe}{\wideas{\dotscale{\Circle}}{\chkYes}}
\newcommand{\chkNo}{\wideas{}{\chkYes}}
\newcommand{\pct}[1]{\SI{#1}{\percent}}
\newcommand{\percentile}[1]{{#1}th~percentile}
\newcommand{\modefont}[1]{\texttt{#1}}
\newcommand{\mnocheck}{\modefont{nocheck}}
\newcommand{\mnonstrict}{\modefont{nonstrict}}
\newcommand{\mstrict}{\modefont{strict}}
\newcommand{\FS}{background}
\newcommand{\zerowidth}[1]{\makebox[0pt][l]{#1}}
\newcommand{\zeroheight}[1]{\raisebox{0pt}[0pt][0pt]{\(#1\)}}
\newcommand{\nK}[1]{#1\textsc{k}}
\newcommand{\stddev}[1]{\nK{#1}}
\newcommand{\gcell}[1]{\cellcolor{green!20}#1}
\newcommand{\ycell}[1]{\cellcolor{yellow!18}#1}
\newcommand{\ocell}[1]{\cellcolor{orange!29}#1}
\newcommand{\rcell}[1]{\cellcolor{red!30}$\!\!$#1$\!\!$}
\newcommand{\gbox}[1]{\colorbox{green!20}{#1}}
\newcommand{\ybox}[1]{\colorbox{yellow!18}{#1}}
\newcommand{\obox}[1]{\colorbox{orange!29}{#1}}
\newcommand{\rbox}[1]{\colorbox{red!30}{#1}}
\newcommand{\tefsfont}[1]{\textsc{#1}}
\newcommand{\tekey}{\tefsfont{te}}
\newcommand{\fskey}{\tefsfont{fs}}
\newcommand{\QALAN}{\FILL{} Alan:}
\newcommand{\panon}{pseudonymized}
\newcommand{\symbox}[1]{{\makebox[0.8em][c]{{#1}}}}
\newcommand{\addsym}{\symbox{\(\uparrow\)}}
\newcommand{\keepsym}{\symbox{\(=\)}}
\newcommand{\dropsym}{\symbox{\(\downarrow\)}}
% \newcommand{\dsymbox}[1]{\makebox[0.8em][c]{#1}}
% \newcommand{\dropsym}{\dsymbox{\raisebox{0ex}[0.9em]{\raisebox{0.5ex}{\vrule width 0.6em height 0.6pt}}}}

@@

\paragraph{Number of Type Analysis Errors}
\label{s:count-analysis-errors}

@@

  \newcommand{\labelbars}[1]{\begin{tabular}[t]{l@{}l} \raisebox{2ex}{\begin{tabular}[t]{r}\mnocheck{}\\\mnonstrict{}\\\mstrict{}\end{tabular}} & #1 \end{tabular}}

    \begin{tabular}[t]{cc}
    %% TODO why text blurry??!
    Type errors & Background errors \\
    \labelbars{\includegraphics[valign=M]{img/error-by-mode-te.pdf}}
    & \labelbars{\includegraphics[valign=M]{img/error-by-mode-fs.pdf}}
  \end{tabular}
  \caption{Type and \FS{} errors grouped by mode.}
  \label{f:error-by-mode}
\end{figure}

Having seen the total number of errors in the dataset and the popularity of the
three analysis modes, a next question is how many errors each mode
accounts for.
There should be a few type errors in the typical \mnocheck{} record (all of
which would be syntax errors) and many more in the \mnonstrict{} and \mstrict{}
records.
For \FS{} errors, we would expect very few in \mstrict{} mode and very many in
\mnocheck{} mode.
Whereas \mnocheck{} mode performs little analysis and therefore gives fews hints
as to how to fix \FS{} errors, \mstrict{} mode checks should closely match
\FS{} checks.

\Cref{f:error-by-mode} divides errors across modes.
By the left plot, \mnonstrict{} accounts for two thirds of all type errors.
This confirms our expectation.
%% because \mnocheck{} reports only syntax errors and
%% \mstrict{} covers only a small number of telemetry records.
The right plot, by contrast, does not support our hypothesis that \mnocheck{}
records have more \FS{} errors than the other modes.
Though \mnocheck{} does have the most \FS{} errors, its proportion matches
the proportion of \mnocheck{} records in the
dataset~(\cref{f:total-records})---if every record had exactly one \FS{} error,
the plot would be the same.

Digging further into \FS{} errors, the median number of such errors by mode is
1 in \mnocheck{}, 3 in \mnonstrict{}, and 1 in \mstrict{}.
Records in \mnocheck{} mode evidently do not have significantly more \FS{}
errors.
Furthermore, type analysis does not seem to reduce \FS{} errors.

@@

\begin{table}[t]
  \caption{Number of telemetry records that increase (\addsym{}), preserve
  (\keepsym{}), or decrease (\dropsym{}) the amount of a specific type
  error from the edit range.}
  \label{t:type-error-survival}

  \begin{tabular}{lr@{~}r@{~}rr@{~}r@{~}rr@{~}r@{~}r}
    & \zerowidth{\mnocheck{}} & & & \zerowidth{\mnonstrict{}} & & & \zerowidth{\mstrict{}} & & \\
    & \rbox{\addsym{}} & \ybox{\keepsym{}} & \gbox{\dropsym{}}
    & \rbox{\addsym{}} & \ybox{\keepsym{}} & \gbox{\dropsym{}}
    & \rbox{\addsym{}} & \ybox{\keepsym{}} & \gbox{\dropsym{}} \\\midrule
    \code{CannotCallNonFunction} & {-} & {-} & {-} & {12} & {2} & {13} & {1} & {-} & {1} \\
    \code{CannotExtendTable} & {-} & {-} & {-} & {6} & {8} & {1} & {-} & {-} & {1} \\
    \code{CannotInferBinaryOperation} & {-} & {-} & {-} & {-} & {1} & {-} & {5} & {6} & {6} \\
    \code{CountMismatch} & {-} & {-} & {-} & {184} & {42} & {157} & {9} & {2} & {9} \\
    \code{DuplicateTypeDefinition} & {-} & {-} & {-} & {1} & {-} & {-} & {-} & {-} & {-} \\
    \code{ExtraInformation} & {-} & {-} & {-} & {46} & {6} & {33} & {1} & {-} & {1} \\
    \code{FunctionDoesNotTakeSelf} & {-} & {-} & {-} & {3} & {4} & {1} & {-} & {-} & {-} \\
    \code{FunctionExitsWithoutReturning} & {-} & {-} & {-} & {6} & {1} & {5} & {7} & {3} & {5} \\
    \code{GenericError} & {-} & {-} & {-} & {177} & {47} & {149} & {5} & {-} & {8} \\
    \code{IllegalRequire} & {-} & {-} & {-} & {7} & {1} & {11} & {-} & {-} & {-} \\
    \code{IncorrectGenericParameterCount} & {-} & {-} & {-} & {-} & {-} & {-} & {1} & {-} & {1} \\
    \code{MissingProperties} & {-} & {-} & {-} & {8} & {3} & {6} & {5} & {3} & {3} \\
    \code{ModuleHasCyclicDependency} & {-} & {-} & {-} & {9} & {5} & {8} & {-} & {-} & {-} \\
    \code{NotATable} & {-} & {-} & {-} & {4} & {1} & {5} & {2} & {-} & {1} \\
    \code{OccursCheckFailed} & {-} & {-} & {-} & {-} & {-} & {-} & {-} & {-} & {1} \\
    \code{OnlyTablesCanHaveMethods} & {-} & {-} & {-} & {-} & {-} & {2} & {-} & {-} & {-} \\
    \code{OptionalValueAccess} & {-} & {-} & {-} & {21} & {45} & {15} & {4} & {2} & {4} \\
    \code{SyntaxError} & {-} & {-} & {8290} & {-} & {-} & {1149} & {-} & {-} & {29} \\
    \code{TypeMismatch} & {-} & {-} & {-} & {103} & {51} & {80} & {13} & {6} & {18} \\
    \code{UnknownPropButFoundLikeProp} & {-} & {-} & {-} & {20} & {17} & {13} & {-} & {-} & {-} \\
    \code{UnknownProperty} & {-} & {-} & {-} & {256} & {156} & {208} & {16} & {18} & {22} \\
    \code{UnknownRequire} & {-} & {-} & {-} & {43} & {30} & {37} & {5} & {3} & {3} \\
    \code{UnknownSymbol} & {-} & {-} & {-} & {1992} & {438} & {1797} & {38} & {18} & {35} \\
  \end{tabular}

@@

The main focus of our telemetry design is to learn how type errors intersect
with edit ranges.
Indeed, most of the fields in a telemetry record are dedicated to this
topic~(\cref{f:tdata}), and it helps to answer RQ2 from the introduction:
\begin{description}
  \item[RQ2.]
    For modules that use type analysis:
    which errors arise,
    how do sessions respond,
    and which errors tend to persist despite subsequent edits?
\end{description}

Using the results of the previous and current type analysis from
each telemetry record,~\cref{t:type-error-survival}
categorizes errors that overlap with the edit range.
The three events of interest are: when edits introduce a type
error, when edits preserve a type error, and when edits remove a
type error.
Hence, the table has three columns for each analysis mode.
In the \mnonstrict{} columns, for example, seventeen records
introduce (\addsym{}) at least one \code{CannotCallNonFunction} error.
Four records keep the number for that error at the same, nonzero level,
and twenty reduce the number of \code{CannotCallNonFunction} errors.

All three counts are imprecise because our telemetry system cannot distinguish
particular type errors.
If a creator removes one \code{NotATable} error but then
immediately introduces another on a different line,
then the centor (\keepsym{}) column gets incremented
instead of the increase (\addsym{}) and decrease (\dropsym{}) columns.

This table is based on the 996,164 records generated from keystroke events.
It ignores records based on module switches because those do not have
meaningful edit ranges.
One implication of the randomized strategy is that we may not notice
when a creator fixes an error; the fix must be selected by the
telemetry system.
%% \QALAN{}: what does an edit range mean for a module-switch record?
%%  there are 2 modules involved but only one range!

@@

\begin{itemize}
  \item
    The numbers in the table are low overall.
    For instance, the highest \mstrict{} count is 38
    records out of the 6,505 total~(\cref{f:dataset-overview}).
    \Cref{s:count-analysis-errors} discussions the implications
    of this low overlap rate for future work.

  \item
    \code{SyntaxError} and the related errors \code{UnknownSymbol} and
    \code{UnknownProperty} are the most popular.
    If the edit range errors are indicative of type errors at large,
    then most errors in the data are likely due to typos.
    These errors also survive (\keepsym{}) the most often.

    \subitem The random-keystroke strategy of our telemetry may be inflating
    the number of typos because it has no way to avoid reporting edits that
    happen in the middle of things.
    For example, a creator who writes a method call
    \code{bucket.countFish()} letter-by-letter will see errors
    including \code{UnknownSymbol(buck)} and \code{UnknownProperty(bucket.c)}.
    Telemetry may pick these up;~\cref{s:conclusion} suggests ways
    to avoid doing so in the future.

  \item
    After the syntax errors, \code{CountMismatch}, \code{TypeMismatch}, and
    \code{GenericError} are the next most common \mnonstrict{} and \mstrict{}
    errors.
    A \code{CountMismatch} sends the wrong number of arguments to a function.
    It may simply be a common error, but it may also be due to our random-keystroke
    telemetry.
    The high incidence of \code{TypeMismatch} and \code{GenericError} calls for
    further study in future work to determine precisely which error arose.
    (Some of the \code{GenericError}s come with context strings in the form of
    \code{ExtraInformation} errors, but telemetry does not report these strings.)
    %% GenericError = only a few things: extend non-class, iterate non-table, ...?

  \item
    \code{OptionalValueAccess} frequently persists through edits (\keepsym{}).
    Creators may be ignoring this error because they can deduce that the
    optional value will be present at runtime.

  \item
    %% Drop >= Add >= Rem ~= 18
    %% other nonzero      ~= 21 triples
    Ideally, the highest number in each column group should
    be the number of decreases (\dropsym{}).
    A decrease should happen when the creator sees an error,
    edits around the highlighted code, and fixes it.
    However, decreases are highest in only 18 of the 39 nonzero groups.

%%  \item
%%    \FILL{} \mnonstrict{} should have no false positives,
%%    \mstrict{} doesn't care.
%%    Why so many keeps in NS??!

%%  \item
%%    A few errors are more common in \mstrict{} than \mnonstrict{}:
%%    \code{CannotInferBinaryOperation},
%%    \code{FunctionExitsWithoutReturn},
%%    \code{IncorrectGenericParamCount}.
%%    \FILL{} so what.

%%   \item
%%     (NOTE may want to save relative popularity for next section.)
%%     Some errors occur more often in \mnonstrict{} than \mstrict{}:
%%     \code{CannotCallNonFunction},
%%     \code{CannotExtendTable},
%%     \code{GenericExtraInformation},
%%     \code{FunctionDoesNotTakeSelf},
%%     \code{IllegalRequire},
%%     \code{ModuleHasCyclicDependency},
%%     \code{OptionalValueAccess},
%%     \code{UnknownPropButGotLikeProp}.
%%     \QALAN{} are any of these surprising, or expected from the bigger number of
%%     \mnonstrict{} records?

@@

@@

\begin{table}[t]\centering
  %% NOTE: popularity looks similar when [over all records] vs. [restricted to module switch records]
  \caption{Type error popularity for \mnonstrict{} and \mstrict{} modes.}
  %% {(all of the type errors in \mnocheck{} mode are \code{SyntaxError}s).}
  \label{t:type-error-count}
  \begin{tabular}{ll}
    \mnonstrict{} & \mstrict{} \\
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{62.13} \\
      \code{SyntaxError} & \pct{15.42} \\
      \code{UnknownProperty} & \pct{8.28} \\
      \code{UnknownRequire} & \pct{3.13} \\
      \code{TypeMismatch} & \pct{2.44} \\
      \code{CountMismatch} & \pct{2.26} \\
      \code{OptionalValueAccess} & \pct{2.07} \\
      \code{GenericError} & \pct{2.06} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.44} \\
      \code{CannotExtendTable} & \pct{0.42} \\
      \code{GenericExtraInformation} & \pct{0.32} \\
      \code{ModuleHasCyclicDependency} & \pct{0.23} \\
      \code{IllegalRequire} & \pct{0.20} \\
      \code{NotATable} & \pct{0.15} \\
      \code{CannotCallNonFunction} & \pct{0.15} \\
      \code{MissingProperties} & \pct{0.09} \\
      \code{FunctionExitsWithoutReturn} & \pct{0.07} \\
      \code{FunctionDoesNotTakeSelf} & \pct{0.07} \\
      \code{MissingUnionProperty} & \pct{0.02} \\
      \code{CannotInferBinaryOperation} & \pct{0.02} \\
      \code{OnlyTablesCanHaveMethods} & \pct{0.01} \\
      \code{DuplicateTypeDefinition} & {<\pct{0.01}} \\
      \code{TypesAreUnrelated} & {<\pct{0.01}} \\
    \end{tabular}
        &
    \begin{tabular}[t]{l@{~}r}
      \code{UnknownSymbol} & \pct{23.97} \\
      \code{TypeMismatch} & \pct{20.46} \\
      \code{UnknownProperty} & \pct{18.88} \\
      \code{SyntaxError} & \pct{9.31} \\
      \code{CannotInferBinaryOperation} & \pct{6.94} \\
      \code{MissingProperties} & \pct{4.04} \\
      \code{CountMismatch} & \pct{2.99} \\
      \code{OptionalValueAccess} & \pct{2.99} \\
      \code{FunctionExitsWithoutReturn} & \pct{2.99} \\
      \code{UnknownRequire} & \pct{2.37} \\
      \code{GenericError} & \pct{2.28} \\
      \code{NotATable} & \pct{1.32} \\
      \code{GenericExtraInformation} & \pct{0.35} \\
      \code{UnknownPropButGotLikeProp} & \pct{0.26} \\
      \code{IncorrectGenericParamCount} & \pct{0.18} \\
      \code{CannotCallNonFunction} & \pct{0.18} \\
      \code{IllegalRequire} & \pct{0.18} \\
      \code{ModuleHasCyclicDependency} & \pct{0.09} \\
      \code{CannotExtendTable} & \pct{0.09} \\
      \code{OccursCheckFailed} & \pct{0.09} \\
      \code{TypesAreUnrelated} & \pct{0.09} \\
    \end{tabular}
  \end{tabular}
\end{table}

@@

\Cref{t:type-error-count} reports the percent of type errors that overlap with
the edit range.
The purpose of this table is to discover all specific errors in the dataset,
thus it includes both keystroke-based and module-switch-based records---even
though edit ranges are not well-defined for the latter.

There are two subtables for \mnonstrict{} and \mstrict{} mode.
The table does not report \mnocheck{} because all its type errors are
by definition \code{SyntaxError}s.

@@

\begin{itemize}
  \item
    Syntax errors, in particular \code{UnknownSymbol}, are much more
    common in \mnonstrict{} than \mstrict{}.
    This may be because \mnonstrict{} reports fewer errors in general.

  \item
    Despite being less strict, \mnonstrict{} reported four errors that
    never appeared in the \mstrict{} records:
    \code{DuplicateTypeDefinition}, \code{OnlyTablesCanHaveMethods},
    \code{MissingUnionProperty}, and \code{FunctionDoesNotTakeSelf}.
    We attribute this to the much larger number of \mnonstrict{} records in the
    dataset.

  \item
    \code{TypeMismatch} is far more common in \mstrict{} (\pct{20.46})
    than \mnonstrict{} (\pct{2.44}) because \mstrict{} mode
    has additional type constraints.
    Several other errors have higher \mstrict{} percentages for the
    same reason, such as \code{CannotInferBinaryOperation}, and
    two errors appeared only in \mstrict{}:
    \code{OccursCheckFailed} and \code{IncorrectGenericParamCount}.
